Iroko: Predictive Datacenter Congestion Control

### 2017-12-10 Documents/Report/intro.tex
More graphs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction / Background}
\label{sec:intro}

@@

Developments in the past decade have changed the general networking 
environment. Data centers have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, data centers must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.
Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
is a perennial challenge for any network operator.
Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the 
eventual collapse of a high-load network, requiring sophisticated 
approaches to solve congestion control.
Despite the innovation potential of data centers, TCP has dominated as 
the congestion control protocol of choice.
TCP is a \textit{reactive protocol}, responding to indicators of congestion and 
latency in the network. However, the fact that packet loss and latency 
surges occur in the network, already portends a problem. Packet loss is 
incurred by overflowing queues in forwarding elements or mismatched 
hardware capabilities, implying that traffic has not been optimally 
distributed. Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

With the advent of Software-Defined Networking (SDN), operators now have the 
ability to freely control and adapt their network architecture, leading to 
highly customized systems and fine-grained optimization.~\cite{sdn_road}

Moving away from the principle of distributed communication and routing, SDN 
introduces the notion of "centralized management". A single controller with 
global knowledge is able to automatically modify and adapt the forwarding 
tables of all switches in the network, while notifying end hosts of changes in 
the network.
These two new trends in system design; full architectural control and 
centralized management, facilitated new opportunities in the space of TCP 
congestion research. Traffic can now be managed in a  \textit{centralized} 
fashion based on \textit{global knowledge} of the entire topology and traffic 
patterns.

A new line of centralized schedulers has emerged that can achieve close to 
optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
However, these schedulers are still  \textit{reactive}  in nature. The central 
controller responds to changes in the network or requests by applications, 
which may cost valuable round-trip latency. Often, short-term flows or bursts 
are unaccounted for, which causes undesirable packet loss and back-propagating 
congestion.

The idea of admission control and service guarantees in networks is not 
new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a datacenter, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behavior, which allows for great simplification of enforcement 
and prioritization policies.

A desirable solution is a global, centralized arbiter which is able to predict 
and fairly distribute flows in the network before bursts or congestion even 
occur. By treating the network's compute and forwarding power as a single 
finite resource, a controller acts akin to a OS scheduler distributing CPU time 
slices to processes. This design approach follows SDN's aspiration of 
introducing operating systems abstractions to the networking domain space.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}

@@

Developments in the past decade have changed the general networking 
environment. Data centers have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, data centers must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.
Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
is a perennial challenge for any network operator.

Inefficient routing and congestion control can quickly lead to 
bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load 
network, requiring more sophisticated solutions. In fact, recent insights into 
datacenter network traffic have shown that buffer overruns and small 
short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc}
Despite the innovation potential of data centers, TCP has dominated as 
the congestion control protocol of choice.
TCP is a \textit{reactive protocol}, responding to indicators of packet drops
latency in the network. However, the fact that loss and latency 
surge in the network, already portends a problem. Overflowing queues in 
forwarding elements or mismatched hardware capabilities imply that traffic has 
not been optimally managed.
Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

Recent research insights and contributions may aid in achieving such a data 
center network.

\paragraph{Software-Defined Networking}
With the advent of Software-Defined Networking (SDN), operators now have the 
ability to freely control and adapt their network architecture, leading to 
highly customized systems and fine-grained optimization.~\cite{sdn_road}

Moving away from the principle of distributed communication and routing, SDN 
introduces the notion of "centralized management". A single controller with 
global knowledge is able to automatically modify and adapt the forwarding 
tables of all switches in the network, while notifying end hosts of changes in 
the network.
This full architectural control and centralized management facilitated new 
opportunities in the space of TCP congestion research. Traffic can now be 
managed in a  \textit{centralized} fashion based on \textit{global knowledge} 
of the entire topology and traffic patterns.

\paragraph{Centralized Scheduling and Congestion Control}
Following the rise of SDN, a new line of centralized schedulers that can 
achieve close to optimal bandwidth utilization has emerged.~\cite{hedera, 
fastpass, microte, b4, dionysus}
However, these schedulers are still \textit{reactive} in nature. The central 
controller responds to changes in the network or requests by applications, 
which may cost valuable round-trip latency. Often, short-term flows or bursts 
are unaccounted for, which causes undesirable packet loss and back-propagating 
congestion.\cite{perc}

\paragraph{Admission-control systems}
In parallel, admission-control-based approaches have gained traction as a 
viable alternative to the conventional burst-and-backoff mechanism of 
TCP.~\cite{expresspass, fastpass, perc}
The idea of admission control and service guarantees in networks is not 
new.~\cite{access_limit, access_limit2}. These designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a datacenter, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behaviour, which allows for great simplification of enforcement 
and prioritization policies. Admission control systems give administrators the 
benefit of tight control over the datacenter, mitigating the effects of 
queue-buildup and bursts.

\paragraph{Predictability of traffic}
Remy\cite{remy} has shown that it is possible to learn optimal congestion 
control parameters for various network topologies, but 
Due the scarcity of public data center traffic sets, it is unclear how 
predictable traffic truly is.

A desirable solution is a global, centralized arbiter, able to predict and 
fairly distribute flows in the network before bursts or congestion even occur. 
By treating the network's compute and forwarding power as a single finite 
resource, a controller acts akin to a OS scheduler distributing CPU time slices 
to processes.

### 2017-12-10 Documents/Report/motivation.tex
More graphs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
\label{sec:motivation}
Sed venenatis odio a augue porttitor, vel blandit tortor luctus. Pellentesque condimentum ante lacus, imperdiet commodo odio mattis a. Donec non consectetur libero. Donec efficitur erat nec dui convallis, sed efficitur nulla pharetra. Nullam sed feugiat est. Vestibulum est nisi, venenatis ac augue molestie, malesuada dignissim dolor. Aenean nisl ante, accumsan nec sagittis quis, euismod eleifend purus. Duis laoreet neque eros, non convallis nunc viverra eget. Sed eleifend neque neque, non faucibus justo vulputate vel. Fusce aliquam ultricies ligula ut efficitur. Nunc ut neque tristique, posuere sem vitae, consectetur nisl.

Donec pretium ut enim ac congue. Suspendisse nisl est, malesuada sit amet posuere a, gravida vitae lacus. Suspendisse et risus mauris. Nam imperdiet sit amet augue sed sodales. Quisque ante orci, aliquet non interdum nec, gravida sed enim. Nunc interdum, purus et scelerisque malesuada, diam lorem efficitur orci, et consequat ligula augue sed lacus. Morbi malesuada eleifend diam eu rhoncus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
\label{sec:motivation}
Several recent contributions in networking research

\paragraph{Software-Defined Networking}
With the advent of Software-Defined Networking (SDN), operators now have the 
ability to freely control and adapt their network architecture, leading to 
highly customized systems and fine-grained optimization.~\cite{sdn_road}

Moving away from the principle of distributed communication and routing, SDN 
introduces the notion of "centralized management". A single controller with 
global knowledge is able to automatically modify and adapt the forwarding 
tables of all switches in the network, while notifying end hosts of changes in 
the network.
This full architectural control and centralized management facilitated new 
opportunities in the space of TCP congestion research. Traffic can now be 
managed in a  \textit{centralized} fashion based on \textit{global knowledge} 
of the entire topology and traffic patterns.

\paragraph{Centralized Scheduling and Congestion Control}
A new line of centralized schedulers has emerged that can achieve close to 
optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
However, these schedulers are still  \textit{reactive}  in nature. The central 
controller responds to changes in the network or requests by applications, 
which may cost valuable round-trip latency. Often, short-term flows or bursts 
are unaccounted for, which causes undesirable packet loss and back-propagating 
congestion.

\paragraph{Admission-control systems}
In parallel, admission-control-based approaches have gained traction as a 
viable alternative to the conventional burst-and-backoff mechanism of 
TCP.~\cite{expresspass, fastpass, perc}
The idea of admission control and service guarantees in networks is not 
new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a datacenter, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behaviour, which allows for great simplification of enforcement 
and prioritization policies.

\paragraph{Predictability of traffic}
